---
title: "Committee Report Fall 2023"
author: "Nick Chang"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# package for data manipulation and visualization
library(tidyverse)
# package to fit generalized linear mixed models
library(glmmTMB)
# package to create interactive plots to more easily explore data
library(plotly)
#package to read in data from excel spreadsheets
library(readxl)
#DHARMa residuals for glmmTMB
library(DHARMa)
#regression tables
library(sjPlot)
# conditional r2
library(performance)

library(car)
#cleaning dates
library(lubridate)

library(MuMIn)

library(lmtest)

#for analyzing compositional data - using aitchison's PCA for substrate
library(compositions)

library(ggbiplot)

```



```{r}

#set wd
setwd("~/Downloads/GitHub/desmog-cheaha-microhabitat-2025/code")

# importing band-level data
bands = read.csv("../data/bands_2025-01-29.csv")
# importing plot-level data
plots = read.csv("../data/plots_2025-01-29.csv")
# importing data about individual organisms
points = read.csv("../data/points_2025-01-29.csv") 
# importing data about capture points (by rock instead of by salamander)
pointsWide = read.csv("../data/pointsWide_2024-12-22.csv")

# Creating a vector of seal salamander streams that we can use to filter later
sealStreams = c("STRM07", "STRM06", "STRM10", "STRM01", "STRM04", "STRM08", "STRM02")


AnonLabels = c(
  "HMR-1" = "Stream 1", 
  "HEAT-3" = "Stream 2", 
  "SHURE-1" = "Stream 3",
  "FOAM-1" = "Stream 4", 
  "VICT-1" = "Stream 5", 
  "DEEP-1" = "Stream 6", 
  "CCC-1" = "Stream 7", 
  "WEEP-1" = "Stream 8", 
  "HAR-1" = "Stream 9", 
  "FERN-1" = "Stream 10", 
  "RICH-2" = "Stream 11"
)

#creating a subset of bands that are from streams that had seal salamanders
ssBands = bands %>% filter(StrmID %in% sealStreams)

#creating a color palette
cols = c('CHOW' = '#0c9896', 
        'PSRUB' = "#C95945",
         "ECIR" = "#E6BC52",
         'EGUT' = '#E68E1C', 
         "DCHE" = "#5D695C", 
          #"DCHE" = "#70665e",
          "DPER" = "#81705c",
          #'DPER' = '#AFA060', 
         'CAM' = "#c7a891", 
        'VAC' = "#90AA86"
         )

colsBi = c('1' = "#5D695C", 
           '0' = "#90AA86")
```

# Introduction

This markdown file contains analyses for my Honors Thesis on the
Talladega seal salamander (*Desmognathus cheaha*) which I have conducted
after consultation with my Thesis committee on November 27th, 2023.




# Microhabitat

At the microhabitat scale, I am interested in what factors predict
microhabitat selection by seal salamanders, particularly whether they
are actively selecting different cover objects.

-   For this dataset, I will be evaluating using binomial
    presence/absence models because while there were a few rocks that
    had multiple seal salamanders under them, that number was very low.
-   I will be filtering my data to only be for streams that have seals
    in them, because I have reason to suspect that there are water
    quality parameters influencing the presence/absence of seal
    salamanders (notably, temperature variation)
-   I will also be filtering my data to only include observations that
    were underneath rocks or wood. Two observations of seal salamanders
    were in burrows, which differ fundamentally in terms of
    microhabitat. In addition, we are dropping observations from
    leafpacks and individuals in the open because there were none of
    these observations with seal salamanders.
-   I am also opting not to include a model that compares habitat use by
    seal salamanders to other species--I am investigating the use of PCA
    to describe habitat use across species re Dr. Martin's suggestion.
-   I will be running logistic regressions where:
    -   1 = seal salamander
    -   0 = no seal salamander BUT may contain other salamanders or
        crayfish

## Microhabitat data transformation

I need to make a few changes to the dataset to perform these analyses.
Below, I am going to:\
\* create a `rocky` substrate composite by adding the percent cover of
gravel and rock. \* create a binary response variable of points that
have seals or are completely vacant (not going to end up using this but
leaving it in the dataset in case it's useful). \* Creating a binary
response variable to compare places with seal salamanders to places with
other salamanders. \* Creating a binary response variable that compares
places that had any salamander to ones that were vacant

In addition, I am creating a subset of the data (`pWS`) which filters
the microhabitat dataset so that my analyses are only include points
from streams where we know that seal salamanders occur. This is because
we believe there are likely some water quality parameters affecting the
ability of other streams to support seal salamanders (see plotwise
models), and that there are likely suitable cover items in these streams
that don't have seals because the habitat isn't suitable for other
reasons.\
I am also filtering to only include observations where the cover type
was rock or wood, because there are some observations of seals in
burrows and I feel that the substrate data isn't comparable.

```{r}

pointsWide = pointsWide %>% 

  #creating a composite of rocky substrate
  mutate(rocky = per_grav + per_rock) %>%

  #creating a response variable for seal vs. fully vacant points
  mutate(sealVac = NA) %>%
  mutate(sealVac = ifelse(Occupied == 0, 0, sealVac)) %>%
  mutate(sealVac = ifelse(DCHE != 0, 1, sealVac)) %>%

  #creating a response variable to compare seals vs. other salamanders
  mutate(sealSal = NA) %>%
  mutate(sealSal = ifelse(salCount > 0, 0, sealSal)) %>%
  mutate(sealSal = ifelse(DCHE != 0, 1, sealSal)) %>%
  
  #creating a response variable for the presence of dusky salamanders
  mutate(duskPres = ifelse(DPER >0, 1, 0)) %>%
  
  #creating a response variable for the presence of 2-lined salamanders
  mutate(cirrPres = ifelse(ECIR > 0, 1, 0 )) %>%
  
  #creating a respnse variable to compare sal vs. non-sal points
  mutate(salVac = ifelse(salCount > 0 & crayCount == 0, 1, 0)) %>%
  
  mutate(nonsealSalCount = SAL + ECIR + DPER +PSRUB + D + EGUT, 
         nonsealOccCount = occupantCount-DCHE) %>%
  mutate(roommate = ifelse(nonsealOccCount >=1, "roommate", "no roommate")) %>%
  mutate(sealStream = ifelse(StrmID %in% c(sealStreams), 1, 0))

```



### Dealing with stream distance

*Sorting things by terrestrial/margin/stream* We are going to create a
categorical variable that sorts salamanders into a
terrestrial/margin/stream variable based on where they were found
in/around the stream. Here, we are going to create some rules that say:

-   Any point with a water depth of 1cm or less AND anything wthin 50cm
    of water is on the margin

-   Anything with a distance to water greater than 50cm is considered
    terrestrial

-   Anything with a deeper water depth is considered in the stream

50cm is considered the cutoff for terrestrial vs. marginal habitat
because when we do our band-level quadrats for marginal habitat we try
to center them on the edge of the stream with 50cm in the stream and
50cm out.

In addition, we are going to create a distDepth variable that combines
the stream distance and water depth variables by subtracting depth from
distance to water so that depth is a negative value.

```{r}
pointsWide = pointsWide %>%
  mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
  mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
  mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
  mutate(hab = as.factor(hab)) %>%
  mutate(distDepth = distWater - cmDepth)
# I am going to create a color palette to make this easier to interpret
habCols = c("stream" = "#9BC4BC", "margin" = "#4B5043", "terrestrial" = "#598218")

```




## Filtering to seal streams


```{r}

#filtering pointsWide to only include points from streams where seal salamanders are present
pWS = pointsWide %>%
  filter(StrmID %in% sealStreams) %>% 
         #filtering to only include observations that were under rock or wood
  filter(CoverType %in% c("R", "W")) 

# pointsWide %>%
#   filter(StrmID %in% sealStreams) %>% 
#   group_by(CoverType) %>%
#   summarize(count = n())


# pWS %>% group_by(sealPres) %>% summarize(count = n())


```




## Fitting a null model

This is how I figured out how to get the marginal/conditional r2: https://github.com/glmmTMB/glmmTMB/issues/169 
```{r}
mic_0 = glmmTMB(sealPres ~ (1|StrmID), data = pWS, family = "binomial")
summary(mic_0)

resid_mic_0 = simulateResiduals(mic_0)
plot(resid_mic_0)

tab_model(mic_0)

confint(mic_0)
summary(mic_0)

exp(-2.1535)

performance::r2(mic_0)

```


# Microhabitat use by Seal Salamanders

## How does particle/pore space and the accumulation of sediment under cover objects affect presence/absence of seal salamanders?

From being in the field, it seems like jumbles of rock tend to make poor
habitat, so I am wondering whether/how that is reflected in the data. I
am also going to separately analyze rock and gravel since it seems like
the difference in particle size would be important. We're including an
interaction parameter for habitat type since there's a thought that
gravelly terrestrial substrates will be less suitable, though the
interaction doesn't seem to be significant.

sealPres \~ (%rock + %gravel) \* hab + (1\|StrmID), data = pWS

```{r}
pWS %>% ggplot(aes(x = rocky, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

```{r}
mic_1.1 = glmmTMB(sealPres ~ rocky + hab + (1|StrmID), data = pWS, family = "binomial")
summary(mic_1.1)
```

The interaction with habitat seems to be very close to 1, so I am also
going to fit a version of the model without the interaction term.

```{r}
mic_1.1 = glmmTMB(sealPres ~ rocky + hab + (1|StrmID), data = pWS, family = "binomial")
summary(mic_1.1)

resid_mic_1.1 = simulateResiduals(mic_1.1)
plot(resid_mic_1.1)

tab_model(mic_1.1)

confint(mic_1.1)



AICc(mic_1.1)

exp( -18.672576852)
```

### Isolating out the gravel variable

When I was doing some of the diagnostic visualizations, it did look a
bit like there was an interesting relationship between seal salamander
presence and the amount of gravel under the rocks.

sealPres \~ (%gravel) * hab + (1\|StrmID)

```{r}
pWS %>% ggplot(aes(x = per_grav, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

```{r}
mic_1.2 = glmmTMB(sealPres ~ per_grav * hab + (1|StrmID)
                    , data = pWS, family = "binomial")
summary(mic_1.2)

resid_mic_1.2 = simulateResiduals(mic_1.2)
plot(resid_mic_1.2)

```

When we compare points where there were seal salamanders or no
salamanders, we do not find a significance.

Dr. Keogh noted that there is lots of random effects variance in this model, so she asked me to pull out StrmID as a fixed effect, which I show here. Deepdene and Fernbank appear to differ from CCC (the stream that the rest are being compared to).

```{r}
mic_1.2.1 = glmmTMB(sealPres ~ per_grav * hab + StrmID
                    , data = pWS, family = "binomial")
summary(mic_1.2.1)

resid_mic_1.2.1 = simulateResiduals(mic_1.2.1)
plot(resid_mic_1.2.1)

```

### Isolating the rock cover variable

I'm also going to have a look at how this plays out when we isolate %
rock as a variable. When I run models with each of these three datasets,
we don't significance in any of them.

sealVac \~ (%rock) + (1\|StrmID)

```{r}
pWS %>% ggplot(aes(x = per_rock, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

When I run this model I get a convergence issue (which is weird because it was working last week), so I need to investigate this further. 

```{r}


mic_1.3 = glmmTMB(sealPres ~ per_rock * hab + (1|StrmID), data = pWS, family = "binomial")
summary(mic_1.3)

resid_mic_1.3 = simulateResiduals(mic_1.3)
plot(resid_mic_1.3)


```



## How does the abundance of leaf litter affect cover object selection for seal salamanders?

I'm interested in whether salamanders are selecting rocks based on the
presence of leaf litter surrounding them. There were a number of rocks
that we found which served as physical infrastructure around which leaf
packs were formed, and presumably the presence of leaf litter allows for
places where salamanders are able to forage and find cover with a ton of
new surface area to chill out in.

sealPres \~ (%litter) + (1\|StrmID)

```{r}
pWS %>% ggplot(aes(x = per_litter, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

```{r}
mic_2.1 = glmmTMB(sealPres ~ per_litter + (1|StrmID), data = pWS, family = "binomial")
summary(mic_2.1)

AICc(mic_2.1)


resid_mic_2.1 = simulateResiduals(mic_2.1)
plot(resid_mic_2.1)

tab_model(mic_2.1)

confint(mic_2.1)

#obtaining the conditional and marginal r2
performance::r2(mic_2.1)

```

## How does the distribution of cover objects of various sizes in and around the stream affect the presence/absence of salamanders?

sealVac \~ coverObjArea \* hab + (1\|StrmID)

hab = stream, margin, terrestrial

For the model comparing seals points to vacant points in the stream, we
see that there is a strong relationship between occupancy and cover
object dimensions, but that there is no significant relationship for
each of the habitat type parameters.

```{r}


# defining the positions that I would like these variables to be in so that they show up correctly int he plots below
positions = c("terrestrial", "margin", "stream")

pWS %>% 
  ggplot() + 
  geom_bar(aes(x = hab, fill = hab)) + 
  scale_fill_manual(values = habCols) + 
  scale_x_discrete(limits = positions) + 
  theme_bw() + 
  
  facet_wrap(~sealVac)


```

```{r}
pWS %>% ggplot(aes(x = covArea, y = sealVac)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

```{r}

mic_3.1 = glmmTMB(sealPres ~ covArea + hab + (1|StrmID), data = pWS, family = "binomial")
summary(mic_3.1)

AICc(mic_3.1)


resid_mic_3.1 = simulateResiduals(mic_3.1)
plot(resid_mic_3.1)


confint(mic_3.1)

tab_model(mic_3.1)

```

### simplifying to only analyze marginal vs. stream habitat
We are thinking that perhaps stream/margin will become significant if we remove terrestrial habitat as a predictor and focus on these two... however as you can see below this is not the case lmao

```{r}

# points2 = points %>% filter(StrmID %in% sealStreams) %>%  
#   mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
#   mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
#   mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
#   mutate(hab = as.factor(hab)) %>%
#   group_by(Species, hab) %>%
#   summarize(x = n())


pWS2 = pWS %>% filter(hab %in% c("stream", "margin"))

mic_3.2 = glmmTMB(sealPres ~ covArea + hab + (1|StrmID), data = pWS2, family = "binomial")
summary(mic_3.2)

```



## How does the presence of salamanders or crayfish affect the cover selection of seal salamanders?

Here, we test whether the presence of other salamanders affects the
presence of seal salamanders under cover items. Since larger cover items
are hypothetically able to support additional individuals, I am going to
include cover area as an interaction effect

```{r}
pWS %>% ggplot(aes(x = nonsealSalCount, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  facet_wrap(~as.factor(nonsealSalCount)) + 
  theme_bw()
```

Another think that I'm thinking about is because larger cover objects are probably able to provide suitable habitat for multiple individual salamanders, how this factors into the suitability of larger rocks. 

```{r}
pWS %>% ggplot(aes(x = covArea, y = sealPres)) + 
  geom_jitter(aes(col = as.factor(nonsealSalCount)), height = .1) + 
  #scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

`roommate` is a categorical variable in which `roommate[roomate]` designates the presence of an additional salamander species or crayfish and `roommate[no roommate]` designates the absence of an additional salamander species or crayfish

```{r}

mic_4.2 = glmmTMB(sealPres ~ roommate * covArea + (1|StrmID), data = pWS, family = "binomial")
summary(mic_4.2)


AICc(mic_4.2)

resid_mic_4.2 = simulateResiduals(mic_4.2)
plot(resid_mic_4.2)

tab_model(mic_4.2)

performance::r2(mic_4.2)

confint(mic_4.2)

fixef(mic_4.2)

shapiro.test(residuals(mic_4.2))


car::Anova(mic_4.2)

simple_4.2 = glm(sealPres ~ roommate * covArea, data = pWS, family = "binomial")

plot(simple_4.2)

```


Obtaining the Odds Ratio estimates for this model

```{r}
#roommate estimate: -1.0320674
exp(-1.0320674)

#cov area estimate:  0.0031035

exp(0.0031035)

#interaction estimate: -0.0018871 
exp(-0.0018871 )

```


```{r}

mic_4.3 = glmmTMB(sealPres ~ roommate + covArea + (1|StrmID), data = pWS, family = "binomial")
summary(mic_4.3)

AICc(mic_4.3)


resid_mic_4.3 = simulateResiduals(mic_4.3)
plot(resid_mic_4.3)

tab_model(mic_4.3)

confint(mic_4.3)

fixef(mic_4.3)

exp(c( -2.422635, -2.233670, 0.001897))


```


```{r}
mic_4.4 = glmmTMB(DPER ~ DCHE, data = pWS, family = "binomial")


```


# FIG 6 LOGISTIC REGRESSION

```{r}




communityFacet = pWS %>% ggplot(aes(x = covArea, y = sealPres
                                  
                                    )) + 
  geom_jitter(height = .05, col = "#5D695C") + 
  #scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", 
              col = "#5D695C",
              method.args = list(family = "binomial"),
              alpha = 0.1) + 
  #labs(title = "Cover Object Area vs. Talladega Seal Salamander Presence/Absence") + 
  xlab("Cover Object Area") + 
  ylab("Seal Salamander Presence/Absence") + 
  theme_bw() + 
  #facet_wrap(~nonsealOccCountBinned, labeller = labeller(nonsealOccCountBinned = as_labeller(occupantLabs))) +
  theme(text = element_text(family = "Times",
    size = 20
    ),
        #legend.position = "none", 
        #removing the ugly gray box behind the facet labels
        strip.background = element_rect(fill = "transparent", color = "black"), 
        strip.placement = "outside", 
        strip.text = element_text(color = "black"),
         panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA))

communityFacet

# ggsave(plot = communityFacet, filename = "../outputs/Figs_HerpConBio/Fig6_LogisticCommunityFacet.pdf")

```

## Posthoc tests for microhabitat models
```{r}

anova(mic_4.2, mic_0, test = "Chisq")


anova(mic_4.2, mic_4.3, test = "Chisq")

```

------------------------------------------------------------------------

# Comparative microhabitat analyses for D. perlapsus

We need to clean up some of the microhabitat data for D perlapsus so that we do not include certain observations (to simplify our levels etc)

We are removing observations which were not from wood or rocks, and we are also filtering to only include microhabitat points from the stream or marginal habitat. 
```{r}
pointsWide %>%
  group_by(as.factor(duskPres), hab) %>%
  dplyr::summarize(n())



duskData = pointsWide %>% filter(CoverType %in% c("W", "R") & hab %in% c("stream", "margin"))

duskData %>% group_by(as.factor(duskPres)) %>% dplyr::summarize(n())

```

Next, we need to join band-level data on cover object density to include this factor as a fixed effect
```{r}
df_covDens = bands %>% select(SubStrmID, covCount, areaSearched) %>%
  mutate(covDens = covCount/areaSearched)


duskData = duskData %>% left_join(df_covDens, by = "SubStrmID")

duskData = duskData %>% mutate(bandNum = str_sub(SubStrmID, -1))

```

## null model
```{r}

dusk_0 = glmmTMB(duskPres ~ + (1|StrmID/bandNum), data = duskData, family = "binomial")


AICc(dusk_0)
```

## habitat type 
```{r}
dusk_1.1 = glmmTMB(duskPres ~  hab*sealStream + covDens + (1|StrmID/bandNum), data = duskData, family = "binomial")
summary(dusk_1.1)


resid_dusk_1.1 = simulateResiduals(dusk_1.1)
plot(resid_dusk_1.1)

AICc(dusk_1.1)
```

## rocky substrate

```{r}
dusk_1.2 = glmmTMB(duskPres ~  rocky*sealStream + covDens + (1|StrmID/bandNum), data = duskData, family = "binomial")
summary(dusk_1.2)

resid_dusk_1.2 = simulateResiduals(dusk_1.2)
plot(resid_dusk_1.2)

AICc(dusk_1.2)
```

## Abundance of leaf litter
```{r}

dusk_2.1 = glmmTMB(duskPres ~ per_litter * sealStream + covDens + (1|StrmID/bandNum), data = duskData, family = "binomial")
summary(dusk_2.1)

resid_dusk_2.1 = simulateResiduals(dusk_2.1)
plot(resid_dusk_2.1)

AICc(dusk_2.1)
```

Presence of leaf litter is not significant

## Distribution of cover objects of various sizes
```{r}

dusk_3.1 = glmmTMB(duskPres ~ covArea*sealStream  + (1|StrmID/bandNum), data = duskData, family = "binomial")
summary(dusk_3.1)

resid_dusk_3.1 = simulateResiduals(dusk_3.1)
plot(resid_dusk_3.1)

AICc(dusk_3.1)
```

## Community dynamics
```{r}
dusk_4.1 = glmmTMB(duskPres ~ covArea + sealPres + (1|StrmID), data = pWS, family = "binomial")
summary(dusk_4.1)


resid_dusk_4.1 = simulateResiduals(dusk_4.1)
plot(resid_dusk_4.1)

AICc(dusk_4.1)
```



