---
title: "Committee Report Fall 2023"
author: "Nick Chang"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#insetwd("/Users/nicholaschang/Downloads/GitHub/desmognathus/code")
# package for data manipulation and visualization
library(tidyverse)
# package to fit generalized linear mixed models
library(glmmTMB)
# package to create interactive plots to more easily explore data
#library(plotly)
#package to read in data from excel spreadsheets
library(readxl)
#DHARMa residuals for glmmTMB
library(DHARMa)

library(plotly)

library(cowplot)

# importing band-level data
bands = read.csv("../data/bands2.csv")
# importing plot-level data
plots = read.csv("../data/plots3_2024-02-27.csv")
# importing data about individual organisms
points = read.csv("../data/points.csv") 
# importing data about capture points (by rock instead of by salamander)
pointsWide = read.csv("../data/pointsWide.csv")

# Creating a vector of seal salamander streams that we can use to filter later
sealStreams = c("CCC-1", "DEEP-1", "FERN-1", "HMR-1", "FOAM-1", "WEEP-1", "HEAT-3")

#creating a subset of bands that are from streams that had seal salamanders
ssBands = bands %>% filter(PlotCode %in% sealStreams)

#creating a color palette
cols = c('CHOW' = '#0c9896', 
        'PSRUB' = "#C95945",
         "ECIR" = "#E6BC52",
         'EGUT' = '#E68E1C', 
         "DCHE" = "#5D695C", 
          #"DCHE" = "#70665e",
          "DPER" = "#81705c",
          #'DPER' = '#AFA060', 
         'CAM' = "#c7a891", 
        'VAC' = "#90AA86"
         )

colsBi = c('1' = "#5D695C", 
           '0' = "#90AA86"
           #'0' = "#2D82B7"
           )

colsTri = c('Seal Salamander' = "#5D695C", 
            "Vacant Point" = "#2D82B7",
           #'Vacant Point' = "#90AA86", 
           'Other Occupant' = "#c7a891")
```

# Introduction

This markdown file contains analyses for my Honors Thesis on the
Talladega seal salamander (*Desmognathus cheaha*) which I have conducted
after consultation with my Thesis committee on November 27th, 2023.

# Microhabitat

At the microhabitat scale, I am interested in what factors predict
microhabitat selection by seal salamanders, particularly whether they
are actively selecting different cover objects.

-   For this dataset, I will be evaluating using binomial
    presence/absence models because while there were a few rocks that
    had multiple seal salamanders under them, that number was very low.
-   I will be filtering my data to only be for streams that have seals
    in them, because I have reason to suspect that there are water
    quality parameters influencing the presence/absence of seal
    salamanders (notably, temperature variation)
-   I will also be filtering my data to only include observations that
    were underneath rocks or wood. Two observations of seal salamanders
    were in burrows, which differ fundamentally in terms of
    microhabitat. In addition, we are dropping observations from
    leafpacks and individuals in the open because there were none of
    these observations with seal salamanders.
-   I am also opting not to include a model that compares habitat use by
    seal salamanders to other species--I am investigating the use of PCA
    to describe habitat use across species re Dr. Martin's suggestion.
-   I will be running logistic regressions where:
    -   1 = seal salamander
    -   0 = no seal salamander BUT may contain other salamanders or
        crayfish

## Microhabitat data transformation

I need to make a few changes to the dataset to perform these analyses.
Below, I am going to:\
\* create a `rocky` substrate composite by adding the percent cover of
gravel and rock. \* create a binary response variable of points that
have seals or are completely vacant (not going to end up using this but
leaving it in the dataset in case it's useful). \* Creating a binary
response variable to compare places with seal salamanders to places with
other salamanders. \* Creating a binary response variable that compares
places that had any salamander to ones that were vacant

In addition, I am creating a subset of the data (`pWS`) which filters
the microhabitat dataset so that my analyses are only include points
from streams where we know that seal salamanders occur. This is because
we believe there are likely some water quality parameters affecting the
ability of other streams to support seal salamanders (see plotwise
models), and that there are likely suitable cover items in these streams
that don't have seals because the habitat isn't suitable for other
reasons.\
I am also filtering to only include observations where the cover type
was rock or wood, because there are some observations of seals in
burrows and I feel that the substrate data isn't comparable.

```{r}

pointsWide = pointsWide %>% 

  #creating a composite of rocky substrate
  mutate(rocky = per_grav + per_rock) %>%

  #creating a response variable for seal vs. fully vacant points
  mutate(sealVac = NA) %>%
  mutate(sealVac = ifelse(Occupied == 0, 0, sealVac)) %>%
  mutate(sealVac = ifelse(DCHE != 0, 1, sealVac)) %>%

  #creating a response variable to compare seals vs. other salamanders
  mutate(sealSal = NA) %>%
  mutate(sealSal = ifelse(salCount > 0, 0, sealSal)) %>%
  mutate(sealSal = ifelse(DCHE != 0, 1, sealSal)) %>%
  
  #creating a respnse variable to compare sal vs. non-sal points
  mutate(salVac = ifelse(salCount > 0 & crayCount == 0, 1, 0)) %>%
  
  mutate(nonsealSalCount = SAL + ECIR + DPER +PSRUB + D + EGUT, 
         nonsealOccCount = occupantCount-DCHE) %>%
  #creating a categorical version of sealVac to use in figure captions
   mutate(sealVacCat = "Other Occupant") %>%
  mutate(sealVacCat= ifelse(Occupied == 0, "Vacant Point", sealVacCat)) %>%
  mutate(sealVacCat = ifelse(DCHE !=0, "Seal Salamander", sealVacCat))

  

#filtering pointsWide to only include points from streams where seal salamanders are present
pWS = pointsWide %>%
  filter(PlotCode %in% sealStreams) %>% 
         #filtering to only include observations that were under rock or wood
  filter(CoverType %in% c("R", "W")) 

```

### Dealing with stream distance

*Sorting things by terrestrial/margin/stream* We are going to create a
categorical variable that sorts salamanders into a
terrestrial/margin/stream variable based on where they were found
in/around the stream. Here, we are going to create some rules that say:

-   Any point with a water depth of 1cm or less AND anything wthin 50cm
    of water is on the margin

-   Anything with a distance to water greater than 50cm is considered
    terrestrial

-   Anything with a deeper water depth is considered in the stream

50cm is considered the cutoff for terrestrial vs. marginal habitat
because when we do our band-level quadrats for marginal habitat we try
to center them on the edge of the stream with 50cm in the stream and
50cm out.

In addition, we are going to create a distDepth variable that combines
the stream distance and water depth variables by subtracting depth from
distance to water so that depth is a negative value.

```{r}
pWS = pWS %>%
  mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
  mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
  mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
  mutate(hab = as.factor(hab)) %>%
  mutate(distDepth = distWater - cmDepth)
# I am going to create a color palette to make this easier to interpret
habCols = c("stream" = "#9BC4BC", "margin" = "#4B5043", "terrestrial" = "#598218")


#sals ISS: Sals in Seal Streams
salsISS = points %>% filter(PlotCode %in% sealStreams) %>%
  mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
  mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
  mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
  mutate(hab = as.factor(hab)) %>%
  mutate(distDepth = distWater - cmDepth)
  

```

## How does particle/pore space and the accumulation of sediment under cover objects affect presence/absence of seal salamanders?

From being in the field, it seems like jumbles of rock tend to make poor
habitat, so I am wondering whether/how that is reflected in the data. I
am also going to separately analyze rock and gravel since it seems like
the difference in particle size would be important. We're including an
interaction parameter for habitat type since there's a thought that
gravelly terrestrial substrates will be less suitable, though the
interaction doesn't seem to be significant.

sealPres \~ (%rock + %gravel) \* hab + (1\|PlotCode), data = pWS

```{r}
rocky = pWS %>% ggplot(aes(x = rocky, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
 stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  labs(title = "Seal salamander presence/absence vs. rocky substrate", 
       subtitle = "Model: Seal Presence/Absence ~ % rocky substrate * habitat type + (1|Stream)", 
       caption = "AIC = 184.6",
       color = "Habitat Type") + 
  ylab("Talladega Seal Salamander\nPresence/Absence") + 
  xlab("Percent cover of rocky substrate\n (% cobble + % gravel)") + 
  theme_bw() 
  #theme(legend.position = "bottom") +

rocky
```


```{r}
mic_1.1 = glmmTMB(sealPres ~ rocky * hab + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_1.1)
```

The interaction with habitat seems to be very close to 1, so I am also
going to fit a version of the model without the interaction term.

```{r}
mic_1.1 = glmmTMB(sealPres ~ rocky + hab + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_1.1)

resid_mic_1.1 = simulateResiduals(mic_1.1)
plot(resid_mic_1.1)
```

### Isolating out the gravel variable

When I was doing some of the diagnostic visualizations, it did look a
bit like there was an interesting relationship between seal salamander
presence and the amount of gravel under the rocks.

sealPres \~ (%gravel) * hab + (1\|PlotCode)

```{r}
pWS %>% ggplot(aes(x = per_grav, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```



### Isolating the rock cover variable



```{r}
pWS %>% ggplot(aes(x = per_rock, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```



## How does the abundance of leaf litter affect cover object selection for seal salamanders?

I'm interested in whether salamanders are selecting rocks based on the
presence of leaf litter surrounding them. There were a number of rocks
that we found which served as physical infrastructure around which leaf
packs were formed, and presumably the presence of leaf litter allows for
places where salamanders are able to forage and find cover with a ton of
new surface area to chill out in.

sealPres \~ (%litter) + (1\|PlotCode)

```{r}
pWS %>% ggplot(aes(x = per_litter, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```



## How does the distribution of cover objects of various sizes in and around the stream affect the presence/absence of salamanders?

sealVac \~ coverObjArea \* hab + (1\|PlotCode)

hab = stream, margin, terrestrial

For the model comparing seals points to vacant points in the stream, we
see that there is a strong relationship between occupancy and cover
object dimensions, but that there is no significant relationship for
each of the habitat type parameters.

```{r}


# defining the positions that I would like these variables to be in so that they show up correctly int he plots below
positions = c("terrestrial", "margin", "stream")

distr = pWS %>% 
  ggplot() + 
  geom_bar(aes(x = hab, fill = hab), show.legend = F) + 
  scale_fill_manual(values = habCols) + 
  scale_x_discrete(limits = positions) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        # Set fill to white and add a black border
        strip.background = element_rect(fill = "white", color = "black"),  
        strip.placement = "outside") + 
  ylab("Count") + 
  xlab("Cover Item Position") + 
  labs(title = "Spatial Distribution of Cover Objects by Occupant Identity") + 
  facet_wrap(~sealVacCat)


distr


pos2 = c("Seal Salamander", "Other Occupant")

distr2 = pWS %>% transform(sealVacCat = factor(sealVacCat, pos2)) %>%
  filter(sealVacCat != "Vacant Point") %>%
  ggplot() + 
  geom_bar(aes(x = hab, fill = hab), show.legend = F) + 
  scale_fill_manual(values = habCols) + 
  scale_x_discrete(limits = positions) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        # Set fill to white and add a black border
        strip.background = element_rect(fill = "white", color = "black"),  
        strip.placement = "outside", 
         plot.title = element_text(size = 18),       # Adjust title font size
  plot.caption = element_text(size = 12)  ) + 
  ylab("Count of Cover Objects") + 
  xlab("Cover Item Position") + 
  labs(title = "Spatial Distribution of Cover Objects used\nby Seal Salamanders")+
  facet_wrap(~sealVacCat)

distr2  + theme( text = element_text(size = 12, family = "Times"))


```


### Visualizations of distribution by species
```{r}

distVis = salsISS %>% 
  mutate(Species = ifelse(Species == "CHOW", "CAM", Species)) %>%
  group_by(Species) %>%
  mutate(spCount = n()) %>%
  ungroup() %>%
  group_by(Species, hab) %>%
  mutate(habProp = n() / spCount) %>%
  ungroup()


salsISS2 = salsISS %>% filter(Species %in% c("DCHE", "DPER"))

  salsISS %>% group_by(Species) %>%
    summarize(n())
```

#FΙGURE 12: STACKED BARPLOTS SHOWING THE DISTRIBUTION ACROSS HABITAT TYPES OF SALAMANDERS AND CRAYFISH BY SPECIES

```{r}


FIG12 = distVis %>% filter(Species %in% c("CAM", "DCHE", "DPER", "ECIR")) %>%
  ggplot(aes(x = Species, y = habProp, fill = factor(hab, levels = c("stream", "margin", "terrestrial")))) +
  geom_bar(stat = "identity", position = "fill",
           aes(col = factor(hab, levels = c("stream", "margin", "terrestrial"))
               )) + 
  scale_fill_manual(values = habCols) + 
    scale_color_manual(values = habCols) +
  guides(color="none") + 
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(y = 1.05, label = paste("n = ", spCount))) + 
  labs(fill = "") + 
  theme_minimal() +
  ylab("Proportion of Observed Animals") + 
  title("Proportions of Commonly Observed Species Observed in Different Habitat Types") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) + theme( text = element_text(size = 12, family = "Times"))

FIG12


```


```{r}
habCov = pWS %>% ggplot(aes(x = covArea, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") +
  labs(title = "Cover Object Area vs. Seal Salamander Presence/Absence", 
       color = " ") + 
  xlab(expression(paste("Cover Object Area (", cm^{2}, ")"))) + 
  ylab("Seal Salamander Presence/Absence") + 
  theme_bw()

habCov
```


```{r}

cov = pWS %>% ggplot(aes(x = covArea, y = sealPres)) + 
  geom_jitter(col = "#5D695C", height = .1) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") +
  labs(title = "Cover Object Area vs. Seal Salamander Presence/Absence", 
       color = " ") + 
  xlab(expression(paste("Cover Object Area (", cm^{2}, ")"))) + 
  ylab("Seal Salamander Presence/Absence") + 
  theme_bw() + theme( text = element_text(size = 14))

cov
```


```{r}

hist(log(pWS$covArea))

x = pWS %>% ggplot(
  aes(label = PointCode, x = as.factor(occupantCount), y = log10(covArea))) + 
    geom_boxplot(fill = "transparent", outlier.shape = NA) + 
  geom_jitter(aes(col = sealVacCat)) + 
  scale_color_manual(values = colsTri) + 
  theme_bw()

ggplotly(x)


pWS = pWS %>% mutate(occCountBinned = ifelse(occupantCount >=2, "2+", occupantCount)) %>%
  mutate(nonsealOccCountBinned = ifelse(nonsealOccCount >=1, "1+", nonsealOccCount))


pWS %>% ggplot(
  aes(label = PointCode, x = as.factor(occCountBinned), y = log10(covArea))) + 
  stat_boxplot(geom ='errorbar', aes(groups = sealVacCat)) + 
  geom_boxplot(#fill = "white", 
    #outlier.shape = NA, 
    aes(fill = sealVacCat,)) + 
  #geom_jitter(aes(col = sealVacCat, groups = sealVacCat)) + 
  scale_fill_manual(values = colsTri) + 
  xlab("Count of Salamanders or Crayfish") + 
  ylab("Cover Object Area (Log 10 Transformed") + 
  theme_bw()
  

pWS %>% 
  ggplot(aes(x = as.factor(occCountBinned), 
             y = log10(covArea))) + 
  geom_boxplot(outlier.shape = NA, 
               position = position_dodge(width = 1), 
               aes(groups = sealVacCat)) + 
  geom_jitter(#position = position_dodge(width = 1),  
              aes(col = sealVacCat)) + 
  scale_color_manual(values = colsTri) + 
  xlab("Count of Salamanders or Crayfish") + 
  ylab("Cover Object Area (Log 10 Transformed") + 
  theme_bw()




box = pWS %>% ggplot(aes(x = reorder(sealVacCat, covArea), y = log(covArea), groups = as.factor(occCountBinned))) + 
    geom_boxplot(#outlier.shape = NA, 
                 position = position_dodge2(width = 0.8, preserve = "single")) + 
  geom_jitter(aes(col = sealVacCat), show.legend = F) + 
  scale_color_manual(values = colsTri) + 
    xlab(" ") + 
  ylab("Cover Object Area\n(log transformed)") + 
  labs(title = "Comparison of Cover Object Area Across Occupied\nand Unoccupied Cover Objects") + 
  theme_bw()
  
  box
  
  
new_labels <- c("Group A" = "Label for Group A", "Group B" = "Label for Group B")
```



```{r}

mic_3.1 = glmmTMB(sealPres ~ covArea + hab + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_3.1)

resid_mic_3.1 = simulateResiduals(mic_3.1)
plot(resid_mic_3.1)

```

```{r}

mic_3.2 = glmmTMB(sealPres ~ hab + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_3.2)

resid_mic_3.1 = simulateResiduals(mic_3.1)
plot(resid_mic_3.1)

```



## How does the presence of salamanders or crayfish affect the cover selection of seal salamanders?


```{r}

occupantLabs = c("0" = "No Additional Community Member", "1+" = "Presence of Add'l Salamander or Crayfish")
  
box2 = pWS %>% 
  ggplot(aes(x = as.factor(sealPres), y = covArea)) + 
    geom_boxplot(outlier.shape = NA
                 #position = position _dodge2(width = 0.8, preserve = "single")
                 ) + 
  geom_jitter(col = "#5D695C", width = 0.1, alpha = 0.5)+ 
 # scale_color_manual(values = colsBi) + 
    xlab("Presence/Absence of Talladega Seal Salamanders") + 
  ylab("Cover Object Area") + 
  labs(title = "Area Comparison of Occupied and Unoccupied\nCover Objects") + 
  theme_bw() + 
  facet_wrap(~nonsealOccCountBinned, labeller = labeller(nonsealOccCountBinned = as_labeller(occupantLabs))) + 
  theme(legend.position = "none", 
        #removing the ugly gray box behind the facet labels
        strip.background = element_rect(fill = "white", color = "black"), 
        strip.placement = "outside")
  
box2 + 
  theme( text = element_text(size = 14, face = "bold"))

  
```

```{R}
box3 = pWS %>% 
  ggplot(aes(x = as.factor(nonsealOccCountBinned), 
             y = covArea, 
             groups = as.factor(nonsealOccCountBinned)
             )
         ) + 
    geom_boxplot(outlier.shape = NA
                 #position = position_dodge2(width = 0.8, preserve = "single")
                 ) + 
  geom_jitter(aes(col = as.factor(sealPres)), 
              #show.legend = F
              ) + 
  scale_color_manual(values = colsBi) + 
    xlab("Presence/Absence of Talladega Seal Salamanders") + 
  ylab("Cover Object Area") + 
  labs(title = "Area Comparison of Occupied and Unoccupied\nCover Objects") + 
  theme_bw() + 
  #facet_wrap(~sealPres) + 
  theme(legend.position = "none", 
        #removing the ugly gray box behind the facet labels
        strip.background = element_rect(fill = "white", color = "black"), 
        strip.placement = "outside")
  
box3 + 
  theme( text = element_text(size = 14, face = "bold"))

  
```

### stacked barplot

```{r}



commVis = pWS %>%
  mutate(sealPresFct = as.factor(sealPres)) %>%
  group_by(nonsealOccCountBinned) %>%
  mutate(cooccCount = n()) %>%
  ungroup() %>%
  group_by(nonsealOccCountBinned, sealPresFct) %>%
  mutate(PresProp = n()/cooccCount) %>%
  ungroup()



commVis2 = commVis %>% group_by(nonsealOccCountBinned, sealPresFct) %>%
  summarize(PresProp = mean(PresProp), count = n())

commVisLab = commVis %>%
  group_by(nonsealOccCountBinned)%>%
  summarize(cooccCount = mean(cooccCount), sealCount = sum(DCHE))


commVisPercents = commVis %>% ggplot(aes(x = nonsealOccCountBinned, y = PresProp, fill = as.factor(sealPres))) + 
  geom_col(position = "fill", aes(col = as.factor(sealPres))) + 
    scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = colsBi) + 
  scale_color_manual(values = colsBi) + 
  guides(color="none") +
  geom_text(aes(y = 1.05, label = paste("n = ", cooccCount))) + 
   # geom_text(aes(y = 0.5, label = paste("n = ", sum(DCHE)))) + 
  theme_minimal()  + 
  labs(fill = "Presence/Absence of \nSeal Salamanders") + 
  ylab("Proportion of Sampled Cover Items") + 
  xlab("Presence/Absence of an Additional Community Member") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  theme( text = element_text(size = 14, face = "bold")) + 
  theme(legend.position = "right")


pWS %>% group_by(as.factor(sealPres), nonsealOccCountBinned)
```

# FIGURE 7: STACKED BARPLOTS SHOWING PROPORTION OF SAMPLED COVER ITEMS THAT HAD SALAMANDERS UNDER THEM IN CO-OCCURRENCE SCENARIOS
```{r}
commVisPercents2 = commVis2 %>% filter(as.factor(sealPresFct) == 1) %>%
  ggplot(aes(x = nonsealOccCountBinned, y = PresProp)) + 
  geom_bar(stat = "identity", fill = "#5D695C") + 
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = colsBi) + 
  scale_color_manual(values = colsBi) + 
  guides(color="none") +
  geom_text(aes(y = PresProp + 0.005, label = paste("n = ", count))) + 
   # geom_text(aes(y = 0.5, label = paste("n = ", sum(DCHE)))) + 
  theme_minimal()  + 
  labs(title =expression(paste("Presence of ", italic("Desmognathus cheaha"), " Across Occupied and Unoccupied Cover")),  
    fill = "Presence/Absence of \nSeal Salamanders") + 
  ylab(expression(paste("Proportion of Sampled Cover Items Sheltering", italic("D. cheaha")))) + 
  xlab("Presence/Absence of an Additional Community Member") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  theme( text = element_text(size = 12, family = "Times")) + 
  theme(legend.position = "right")

commVisPercents2


```


```{r}
commVis %>% ggplot(aes(x = nonsealOccCountBinned, fill = as.factor(sealPres))) + 
  geom_bar( aes(col = as.factor(sealPres))) + 
  scale_fill_manual(values = colsBi) + 
  scale_color_manual(values = colsBi) + 
  guides(color="none") +
  #geom_text(aes(y = cooccCount + 3, label = paste("n = ", cooccCount))) + 
   # geom_text(aes(y = 0.5, label = paste("n = ", sum(DCHE)))) + 
  theme_minimal()  + 
  labs(fill = "Presence/Absence of \nSeal Salamanders") + 
  ylab("Count of Sampled Cover Items") + 
  xlab("Presence/Absence of an Additional Community Member") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  theme( text = element_text(size = 14, face = "bold")) + 
  theme(legend.position = "right")

```








<<<<<<< Updated upstream
FIGURE 6: LOGISTIC REGRESSION SHOWING SEAL/PRESENCE ABSENCE AS A FUNCTION OF COVER OBJECT AREA, FACETED BY CO-OCCURRENCE OF COMMUNITY MEMBER
=======
Community Facet
>>>>>>> Stashed changes

```{r}

pWS = pWS %>% mutate(roommate = ifelse(nonsealOccCountBinned == "0", 0, 1))


communityFacet = pWS %>% ggplot(aes(x = covArea, y = sealPres
                                  
                                    )) + 
  geom_jitter(height = .05, col = "#5D695C") + 
  #scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", 
              col = "#5D695C",
              method.args = list(family = "binomial"),
              alpha = 0.1) + 
  #labs(title = "Cover Object Area vs. Talladega Seal Salamander Presence/Absence") + 
  xlab("Cover Object Area") + 
  ylab("Seal Salamander Presence/Absence") + 
  theme_bw() + 
  #facet_wrap(~nonsealOccCountBinned, labeller = labeller(nonsealOccCountBinned = as_labeller(occupantLabs))) +
  theme(text = element_text(family = "Times",
    size = 20
    ),
        #legend.position = "none", 
        #removing the ugly gray box behind the facet labels
        strip.background = element_rect(fill = "transparent", color = "black"), 
        strip.placement = "outside", 
        strip.text = element_text(color = "black"),
         panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA))

communityFacet

# ggsave(plot = communityFacet, filename = "../outputs/Figs_HerpConBio/Fig6_LogisticCommunityFacet.pdf")

```


```{r}

pWS %>% filter(sealPres == 1) %>% ggplot() + 
  geom_boxplot(aes(x = as.factor(roommate), y = covArea)) +
  geom_jitter(aes(x = as.factor(roommate), y = covArea)) 


pWS %>% filter(sealPres==1) %>% ggplot() + 
  geom_density(aes(x = covArea, fill = as.factor(roommate)), alpha = 0.5)


pWS %>% ggplot() + 
  geom_density(aes(x = covArea, fill = as.factor(sealPres)), alpha = 0.5) + 
  facet_wrap(~as.factor(roommate)) + theme_bw()

```

```{r}

mic_4.1 = glmmTMB(sealPres ~ nonsealSalCount * covArea + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_4.1)

resid_mic_4.1 = simulateResiduals(mic_4.1)
plot(resid_mic_4.1)

```

Here, we expand the above analysis to include crayfish and nonseal
salamanders

```{r}
pWS %>% ggplot(aes(x = nonsealOccCount, y = sealPres)) + 
  geom_jitter(aes(col = hab), height = .1) + 
  scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

When we include crayfish in this analysis, we can see that nearly every cover object with an area greater than 1000 cm^2 has some occupant. This is reflected in the model below, which finds a significant interaction between count of non-seal salamander occupants and the size of the cover item. 


```{r}
pWS %>% ggplot(aes(x = covArea, y = sealPres)) + 
  geom_jitter(aes(col = as.factor(nonsealOccCount)), height = .1) + 
  #scale_color_manual(values = habCols) + 
  stat_smooth(method = "glm", method.args = list(family = "binomial")) + 
  theme_bw()
```

```{r}
pWS %>% ggplot() + 
  geom_bar(aes(x= nonsealOccCount))


pWS %>% group_by(nonsealOccCount) %>% summarize(count = n())




mic_4.2 = glmmTMB(sealPres ~ nonsealOccCount * covArea + (1|PlotCode), data = pWS, family = "binomial")
summary(mic_4.2)

resid_mic_4.2 = simulateResiduals(mic_4.2)
plot(resid_mic_4.2)


```


## Cover Type
```{r}
points %>% filter(Species == "DCHE") %>%
  ggplot() + 
  geom_bar(aes(x = factor(CoverType, levels = c("R", "B", "W"))), fill = "#5D695C") +
  theme_minimal() + 
  ylab("Count") + 
  xlab("Cover Item Identity")

```

------------------------------------------------------------------------

# Plot Figures

I am working with the following predictor variables at the stream level: 


* Daily mean stream temperature
* Dail mean stream temperature amplitude
* Dissolved oxygen
* water depth
* channel height
* stream declivity
* watershed area
* % imperv

```{r}
plots = plots %>% mutate(sealStreamLabel = ifelse(sealStream == 1, "Seal Salamanders \n Detected", "Seal Salamanders \n Not Detected"))



```


## Daily mean stream temperature

```{r}

meantemp = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = dailyMeanTemp)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Daily Mean Temperature (°C)") + 
  ylab("") + 
  theme_bw()

meantemp


```

## Daily temp amplitude

```{r}
tempAmp = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = dailyMeanAmpl)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
#  labs(title = "Temperature variation in streams with and without Seal Salamanders") +
  xlab("Daily Mean Temp. Amplitude (°C)") + 
  ylab("") + 
  theme_bw()

tempAmp

```


## DO

```{r}
DOplot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = DO)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
  #labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Dissolved Oxygen (%)") + 
  ylab("") + 
  theme_bw()

DOplot

```



## depth

```{r}
depthPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = depth)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
  #labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Mean Water Depth (cm)") + 
  ylab("") + 
  theme_bw()

depthPlot

```




## channel height

```{r}
cHeiPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = cHei)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Mean Channel Height (cm)") + 
  ylab("") + 
  theme_bw()

cHeiPlot


```


## decliv

```{r}
declivPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = Declivity)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
  #labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Declivity (°)") + 
  ylab("") + 
  theme_bw()

declivPlot


```


## watershed area

```{r}
wsaPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = DrainageArea)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
  #labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab(expression(paste("Watershed Area (", m^{2}, ")"))) + 
  ylab("") + 
  theme_bw()

wsaPlot


```


## imperv

```{r}
impervPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = imperv)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Impervious Surface Cover (%)") + 
  ylab("") + 
  theme_bw()

impervPlot


```

## cover

```{r}
plots = plots %>% mutate(covDens = covCount/areaSearched)
covPlot = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = covDens)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab(expression(paste("Cover Item Density (objects/", m^{2}, ")"))) + 
  ylab("") + 
  theme_bw()

impervPlot


```

## dbh

```{r}
dbh = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = meanDBH)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Mean DBH of Surrounding Trees (cm)") + 
  ylab("") + 
  theme_bw()

dbh

```

```{r}
maxdbh = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = maxDBH)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab("Mean DBH of Surrounding Trees (cm)") + 
  ylab("") + 
  theme_bw()

maxdbh

dbh2 = glm(sealStream ~ meanDBH, data = plots, family = "binomial")
summary(dbh)

```

## conductivity
```{r}
conduct = ggplot(data = plots, aes(y = sealStreamLabel, 
                   x = Conduct)) + 
    geom_boxplot(outlier.shape = NA, 
               fill = "transparent") + 
  geom_jitter(size = 3, height = 0.1, 
               col = "#5D695C") + 
 # labs(title = "Comparison of temperature variaton in streams with and without seal salamanders") +
  xlab(expression(paste("Conductivity (", mu, "S/cm)"))) + 
  ylab("") + 
  theme_bw()

conduct

cd = glm(sealStream ~ Conduct, data = plots, family = "binomial")

AICc(cd)
summary(cd)

tab_model(cd)
```



```{r}

plot_grid(meantemp,  tempAmp,  declivPlot, DOplot, conduct, depthPlot, cHeiPlot, covPlot, dbh, wsaPlot, impervPlot, ncol = 2, scale = 0.9)

```

```{r}
plots %>% ggplot(aes(x =Declivity, y = PlotCode)) + 
  geom_bar(stat = "identity")

glimpse(plots)

width = glm(sealStream ~ cWid, data = plots, family = "binomial")
summary(width)

```






# OLD PLOT FIGURES

## How do watershed-influences of urbanization affect seal salamander presence and density?

I am interested in how watershed-level characteristics affect seal
salamander density. I used ArcGIS to calculate the percent coverage of
impervious surface upstream from my plots and incorporated that here.

The other variable that I am using here is daily temperature variation,
which I believe serves as a proxy for the realized effect of impervious
surface cover on the landscape given the way that pavement leads to heat
islands, and how water flowing over pavement and into streams will
absorb heat. I am hoping that greater fluctuations will also represent
pulses of water through the landscape.

```{r, message = F}

ggplot(plots, aes(x = imperv, y = sealStream)) +
  geom_hline(yintercept = 0)+
    stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  #geom_smooth(method = "lm", alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  geom_point(aes(col = imperv), size = 4) + 
 # ylim(-.1, 0.6) +
  theme_bw() + 
   # theme(legend.position = "bottom") + 
  labs(title = "Seal salamander density vs. impervious surface cover") + 
  xlab("Watershed Imperviousness (%)") + 
  ylab("Seal salamander presence")
  #ylab(expression(paste("Talladega Seal Salamander Density (", m^{2}, ")")))

```

```{r, message = F}

# ggplot(plots, aes(x = dailyMeanAmpl, y = DCHE_m2)) +
#   geom_hline(yintercept = 0)+
#   geom_smooth(method = "lm", alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
#   geom_point(aes(col = dailyMeanAmpl), size = 4) + 
#   scale_color_gradient(low = "darkblue"#"#202A44"
#                           , high = "orange" ) + 
# 
#  # ylim(-.1, 0.6) +
#   theme_bw() + 
#   #  theme(legend.position = "bottom") + 
#   labs(title = "Seal salamander density vs. temperature variation", color = " ") + 
#   xlab("Average daily stream temperature range (°C)") + 
#   ylab(expression(paste("Talladega Seal Salamander Density (", m^{2}, ")")))



#this one is a logistic one


x = ggplot(plots, aes(x = dailyMeanAmpl, y = sealStream)) +
    geom_hline(yintercept = 0)+
  geom_point(aes(col = dailyMeanTemp), size = 3) +
   scale_color_gradient(low = "darkblue"#"#202A44"
                           , high = "orange" ) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  theme_bw() +
  theme(legend.position = c(0.165, 0.3),
    legend.background = element_rect(fill = "white", color = "black")) + 
 # theme(legend.position = "bottom") + 
  labs(title = "Salamander Presence/Absence vs. Temperature Variation", 
       color = "Daily Mean\nTemperature (°C)")+ 
  ylab("Talladega Seal Salamander Presence") + 
  xlab("Daily Mean Stream Temperature Amplitude (°C)")


x

x + theme(
  text = element_text(size = 14, face = "bold")
)



```

```{r}
ggplot(data = plots) + 
    geom_boxplot(outlier.shape = NA, 
               aes(x = as.factor(sealStream), 
                   y = avgDiVar), fill = "transparent") + 
  geom_jitter(size = 3, width = 0.25, 
              aes(x = as.factor(sealStream), 
                  y = avgDiVar, 
                  col = as.factor(sealStream))) + 
  scale_color_manual(values = colsBi) + 
  labs(title = "Comparison of temperature variaton in streams with and without seal salamanders", 
       col = "Seal presence") +
  theme_bw()

```

DCHE \~ Temperature variation + imperviousness

```{r}

plots_1.1 = glmmTMB(sealStream ~ dailyMeanAmpl + imperv, family = "binomial", data = plots)
summary(plots_1.1)


resid_plots_1.1= simulateResiduals(plots_1.1)
plot(resid_plots_1.1)
```

When we fit this model including temperature variation and watreshed
imperviousness, we get a positive estimate for the effect of watershed
imperviousness. I think this is likely because of site CCC in Wesley
Woods which is piped under the conference center and has a lot of seal
salamanders. From conversations with John, construction here was pretty
good about stormwater infrastructure so I am not sure how I should feel
about this.



Below is a model that just has temperature variation.

```{r}
plots_1.1.1 =  glmmTMB(sealStream ~ dailyMeanAmpl,family = "binomial",  data = plots)
summary(plots_1.1.1)

resid_plots_1.1.1 = simulateResiduals(plots_1.1.1)
plot(resid_plots_1.1.1)
```

```{r}
plots_1.1.2 =  glmmTMB(sealStream ~ dailyMeanTemp,family = "binomial",  data = plots)
summary(plots_1.1.2)

resid_plots_1.1.1 = simulateResiduals(plots_1.1.2)
plot(resid_plots_1.1.2)
```



## Landcover dataset

I was able to extract data from the NLCD Landfire analysis for each
watershed, but I am not sure at this point what exactly to analyze.
However, I do at least want to have a look at whether watershed area is
a relevant predictor since it seems like that may be relevant to the
lack of seals in some of the higher order streams.

```{r}
plots = plots %>%
  mutate(perDev = devLow_perc + devMed_perc + devOpen_perc) %>%
  mutate(perDist = perDev + dryLob_perc + field_perc + pine_perc, 
         sumDist = sumDev + dryLob_sum + field_sum + pine_sum, 
         perMature = mesic_perc + dryHard_perc, 
         sumMature = mesic_sum + dryHard_sum, 
         perForest = dryLob_perc + pine_perc + mesic_perc + dryHard_perc + dryMixed_perc, 
         sumForest = dryLob_sum + pine_sum + mesic_sum + dryHard_sum + dryMixed_sum, 
         watershedArea = dryLob_sum + pine_sum + mesic_sum + dryHard_sum + dryMixed_sum + field_sum + devLow_sum + devMed_sum + devOpen_sum
           )
```




```{r}

ggplot(plots, aes(x = watershedArea, y = sealStream)) +
  geom_hline(yintercept = 0)+
   stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  geom_point( size = 4) + 
 # ylim(-.1, 0.6) +
  theme_bw() + 
    #theme(legend.position = "bottom") + 
  labs(title = "Seal salamander presence vs. Watershed Area") + 
  xlab(expression(paste("Watershed Area (", m^{2}, ")"))) + 
  ylab("seal salamander presence")
  #ylab(expression(paste("Talladega Seal Salamander Density (", m^{2}, ")")))


plots_4.1 = glmmTMB(sealStream ~ watershedArea, data = plots, family = "binomial")
summary(plots_4.1)

resid_plots_4.1 = simulateResiduals(plots_4.1)
plot(resid_plots_4.1)

```


```{r}
plots_4.2 = glmmTMB(sealStream ~sumDev, data = plots, family = "binomial")
summary(plots_4.2)

resid_plots_4.2 = simulateResiduals(plots_4.2)
plot(resid_plots_4.2)


ggplot(plots, aes(x = sumDev, y = dailyMeanAmpl)) + 
  geom_point() + 
  geom_smooth(method = "lm")


devArea = ggplot(plots, aes(x = sumDev, y = sealStream)) +
  geom_hline(yintercept = 0)+
   stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  geom_point( size = 4, aes(col = as.factor(sealStream))) + 
  scale_color_manual(values = colsBi, guide = "none") +
 # ylim(-.1, 0.6) +
  theme_bw() + 
    #theme(legend.position = "bottom") + 
  labs(title = "Salamander Presence/Absence vs. Developed Area") + 
  xlab(expression(paste("Developed Area Upstream in the Watershed (", m^{2}, ")"))) + 
  ylab("Talladega Seal Salamander Presence") + 
  theme( text = element_text(size = 14, face = "bold"))
  
  #ylab(expression(paste("Talladega Seal Salamander Density (", m^{2}, ")")))




grid = plot_grid(x, devArea, align = "v")
```

## What influence does stream-level geomorphology have on salamander presence?

Dr. Beck suggested checking to see if band-level variables were
spatially autocorrelated, so I plotted them out below. It does appear
that way, so I have incorporated the stream geomorphology

```{r}

bands %>%
ggplot(aes(x = bandNum, y = cHei)) + 
  geom_bar(aes(fill = StreamID), stat = "identity") + 
  theme_bw() + 
  facet_wrap(~StreamID) + 
  labs(title = "Channel height across sites")

bands %>%
ggplot(aes(x = bandNum, y = depth)) + 
  geom_bar(aes(fill = StreamID), stat = "identity") + 
  theme_bw() + 
  facet_wrap(~StreamID) + 
  labs(title = "water depth across sites")

```

Channel height appears to be relatively important in understanding seal
sal density at each of these streams

```{r}
plots %>% ggplot(aes(x = cHei, y = sealStream)) + 
      geom_hline(yintercept = 0)+
    stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              alpha = 0.10, col = "#5D695C", fill = "#5D695C") + 
  geom_point(size = 4, col = "#5D695C") + 
  xlab("Channel Height")+ 
  ylab("Seal salamander presence") + 
  theme_bw()


plots_3.1 = glmmTMB(sealStream ~ cHei, family = "binomial",  data = plots)
summary(plots_3.1)
```


```{r}
plots_5.1 = glmmTMB(sealStream ~ covCount, family = "binomial",  data = plots)
summary(plots_5.1)

```

## How do turbidity and conductivity correlate with salamander density?

I am not analyzing turbidity data right now because I think that I was
not fast enough to collect all of my samples in time for samples to
accurately reflect differences in turbidity between sites. As a part of
this, I have two conductivity datasets (one reflecting when I sampled
each stream and one relflecting when I took turbidity samples).

We can see that the correlation between these is not that strong...
which is interesting.

```{r}
plots%>% select(altCond, Conduct) %>% cor()

plots %>% ggplot(aes(x = altCond, y = Conduct)) + geom_smooth(method = "lm") + geom_point() + theme_bw()

```
