---
title: "Community Dynamics"
author: "Nick Chang"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(glmmTMB)
library(car)
library(sjPlot)
library(performance)
library(ggeffects)
library(MuMIn)
library(lme4)
library(readxl)

```


Reading in data
```{r}
#pointsPlots = read.csv('../data/pointsPlots.csv')

points = read.csv("../data/points_2025-01-29.csv")

plots = read.csv("../data/plots_2025-01-29.csv")

pointsWide = read.csv("../data/pointsWide_2024-12-22.csv")

bands = read.csv("../data/bands_2025-01-29.csv")

bandRocks = read.csv("../data/MicrohabAvail_2025-02-20.csv")


cols = c('CHOW' = '#0c9896', 
        'PSRUB' = "#C95945",
         "ECIR" = "#E6BC52",
         'EGUT' = '#E68E1C', 
         "DCHE" = "#5D695C", 
          #"DCHE" = "#70665e",
          "DPER" = "#81705c",
          #'DPER' = '#AFA060', 
         'CAM' = "#c7a891"
         )

habCols = c("stream" = "#9BC4BC", "margin" = "#4B5043", "terrestrial" = "#598218")

```


Reading out a breakdown of how many escaped, NOID desmognathus there were: 
```{r}


points %>% group_by(Species, Escaped) %>% summarize(n())

points %>% filter(Species %in% c("D", "SAL")) %>% select(Escaped, Species)



pointsWide %>% filter(DCHE > 1) %>% select(CoverType, DCHE)

```


Creating a plot to show the size class distribution of D perlapsus based on whether they were in streams with seal salamanders
```{r}
points %>%
  filter(Species %in% c('DPER')) %>%
  ggplot(aes(x = mmSVL)) + 
  geom_histogram(aes(fill = as.factor(sealStream)), alpha = 0.5, position = "identity", bins = 20) + 
 # facet_wrap(~sealStream.x) + 
  theme_bw()

```



# Stream distance/cover position

```{r}
#creating the cover item position variable (stream/margin/terr)
points = points  %>% 
  mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
  mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
  mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
  mutate(hab = as.factor(hab)) %>%
  mutate(distDepth = distWater - cmDepth)
# I am going to create a color palette to make this easier to interpret

```

# Do seals utilize more aquatic habitat than other species

```{r}

points %>% filter(sealStream==1) %>% 
  mutate(Species = ifelse(Species == "CHOW", "CAM", Species)) %>% 
    filter(Species %in% c("CAM", "DCHE", "DPER", "ECIR")) %>% group_by(Species) %>% summarize(count = n())


pointsSealStream = points %>% filter(sealStream == 1) %>% filter(hab != "terrestrial") %>% 
  mutate(Species = ifelse(Species == "CHOW", "CAM", Species)) %>% filter(Species %in% c("CAM", "DCHE", "DPER", "ECIR")) %>% 
  mutate(Species = ifelse(Species == "CAM", "xCAM", Species))

pointsSealStream %>% group_by(Species) %>% summarize(count = n())


```


```{r}
pointsSealStream = pointsSealStream %>% mutate(habBin = ifelse(hab == "stream", 1, 0))

nullComp = compHab = glmmTMB(habBin ~ (1|StrmID), data = pointsSealStream, family = "binomial")


compHab = glmmTMB(habBin ~ Species + (1|StrmID), data = pointsSealStream, family = "binomial")
summary(compHab)

tab_model(compHab)


anova(nullComp, compHab, test = "Chisq")

exp(-0.32)
```

## visualization

```{r}

dist = pointsSealStream %>%
  group_by(Species) %>%
  mutate(spCount = n()) %>%
  ungroup() %>%
  group_by(Species, hab) %>%
  mutate(habProp = n() / spCount) %>%
  ungroup()


FIG12 = dist %>% filter(Species %in% c("CAM", "DCHE", "DPER", "ECIR")) %>%
  ggplot(aes(x = Species, y = habProp, fill = factor(hab, levels = c("stream", "margin", "terrestrial")))) +
  geom_bar(stat = "identity", position = "fill",
           aes(col = factor(hab, levels = c("stream", "margin", "terrestrial"))
               )) + 
  scale_fill_manual(values = habCols) + 
    scale_color_manual(values = habCols) +
  guides(color="none") + 
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(y = 1.05, label = paste("n = ", spCount))) + 
  labs(fill = "") + 
  theme_minimal() +
  ylab("Proportion of Observed Animals") + 
  #title("Proportions of Commonly Observed Species Observed in Different Habitat Types") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) + theme( text = element_text(size = 12, family = "Times"))

FIG12


```

# Is the proportion of duskies using stream/marginal habitat different depending on whether there are seals in the stream?
```{r}
points %>% filter(Species == "DPER") %>% 
  ggplot() + 
  geom_bar(aes(x = hab)) + 
  facet_wrap(~sealStream)


dper = points %>% filter(Species %in% c("DPER")) %>%
  group_by(sealStream, Species) %>%
  dplyr::mutate(sSCount = dplyr::n()) %>%
  ungroup() %>%
  group_by(sealStream, hab, Species) %>%
  dplyr::mutate(habProp = dplyr::n() / sSCount) %>%
  ungroup()


```

# FIGURE 11: STACKED BARPLOT SHOWING DISTRIBUTION OF D PERLAPSUS

```{r}


dper2 = dper %>% filter(hab %in% c("stream", "margin"))

fig11 = dper2 %>%
  ggplot(aes(x = as.factor(sealStream), y = habProp, fill = factor(hab, levels = c("stream", "margin", "terrestrial")))) +
  geom_bar(stat = "identity", position = "fill",
           aes(col = factor(hab, levels = c("stream", "margin", "terrestrial"))
               )) + 
  scale_fill_manual(values = habCols) + 
    scale_color_manual(values = habCols) +
  guides(color="none") + 
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(y = 1.05, label = paste("n = ", sSCount))) + 
  labs(fill = ""#, 
      # title = expression(paste("Cover Item Positions of ", italic("Desmognathus perlapsus")))
       ) + 
  theme_minimal() +
  ylab("Proportion of Observed Animals") + 
  xlab("Detection of Seal Salamanders in the Sampled Stream") + 
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    text = element_text(size = 20, family = "Times")
  ) + 
  coord_flip()

fig11

```

# Dusky stream habitat use


First, I need to clean data

I need to add reach-level 

```{r}
#bandRocks = read_excel("../data/PlotData_Spring2023.xlsx", sheet = "MicrohabitatAvailability2")


```


### are there differing counts of rocks across streams?

```{r}

rockMod = lm(rockCount ~ StrmID, data = bandRocks)


summary(rockMod)

x = Anova(rockMod)

summary(x)


```

```{r}

bandRocks = bandRocks %>% mutate(SubplotHab = paste(SubplotCode, HabitatType)) %>% select(SubplotHab, rockCount)

dper = dper %>% separate(SubplotCode, into = c("Stream2", "plot", "band"), sep = "-")

dper = dper %>% mutate(SubplotHab = paste0(StrmID, "-", band, " ", toupper(hab)))

dper = dper %>% filter(hab %in% c("margin", "stream")) %>% left_join(bandRocks, by = "SubplotHab")

dper = dper %>% filter(duplicated(PointCode) == FALSE)

```





```{r}

dper2 = dper %>% filter(hab %in% c("stream", "margin") & CoverType %in% c("R", "W")) %>%
  mutate(hab2 = ifelse(hab == "stream", 1, 0))

dper2 %>% ggplot() + 
  geom_bar(aes(x = CoverType))

glimpse(dper2)





# summary(dperMod)
#   
# tab_model(dperMod)
# 
# confint(dperMod)

exp(-1.0884824)
```



Testing whether habitat utilization is dependent on the presence of 
D. cheaha in the stream
```{r}
dperNull = glmmTMB(hab2 ~ (1|StrmID), data = dper2, family = "binomial")

dperMod = glmmTMB(hab2 ~ sealStream + rockCount + (1|StrmID), data = dper2, family = "binomial")

dperSimpl = glmmTMB(hab2 ~ sealStream + (1|StrmID), data = dper2, family = "binomial")

AICc(dperMod)
AICc(dperSimpl)

anova(dperMod, dperNull, test = "Chisq")
anova(dperMod, dperSimpl, test = "Chisq")




  summary(dperMod)
summary(dperSimpl)

tab_model(dperMod)

# r.squaredGLMM(dperMod)
# 
 r2_nakagawa(dperMod, tolerance = 0)

exp(-0.2153)

confint(dperMod)



```


## Manually calculating marginal and conditional R2
```{r}
# mF = glmer(hab2 ~ sealStream + covDens + (1|PlotCode/bandNum), data = dper3, family = "binomial")


# r.squaredGLMM(mF)
# 
# 
# 
# summary(mF)
summary(dperMod)

# Extraction of fitted value for the alternative model
# fixef() extracts coefficents for fixed effects
# mF@X returns fixed effect design matrix
# Fixed <- fixef(mF)[2] * mF@X[, 2] + fixef(mF)[3] * mF@X[, 3] + fixef(mF)[4] * mF@X[, 4]
# 
# # Calculation of the variance in fitted values
# VarF <- var(Fixed)
# 
# # An alternative way for getting the same result
# VarF <- var(as.vector(fixef(mF) %*% t(mF@X)))

```



```{R}

summary(dperMod)

tab_model(dperMod)

confint(dperMod)

r2(dperMod)

check_singularity(dperMod)


```




# Morphometrics

## Sample Size Overview
```{r}
points %>% filter(Occupied == 1) %>%
  group_by(Species, Escaped) %>%
  summarize(n())
```

Across all of our study sites, we had 24 captures of 
D. cheaha and 10 escapes. 

```{r}

points %>% filter(Occupied == 1 & Species == "DCHE") %>%
  group_by(StrmID, Escaped) %>%
  summarize(n())

```

When we look at a breakdown of escapes at each site, we find that we did not have any successful captures of D cheaha, at two of our sites which leaves our count of streams with successful captures at n = 5. 

```{r}
points %>% filter(Occupied ==1 & Species == "DCHE" & Escaped == "NO") %>%
  group_by(StrmID)%>%
  summarize(n())
```


```{r}
pointsDCHE = points %>% filter(Species == "DCHE")

pointsDCHE2 = pointsDCHE %>% filter(Escaped =="NO" & CoverType %in% c("R", "W")) %>%
  mutate(rocky = per_grav + per_rock)



```


## Which size metrics to look at?
I have multiple size metrics: 
 * Snout-vent length
 * total length
 * mass in grams
 * tail score



I expect that larger seal salamanders will occupy larger cover objects


## habitat changes and salamander size
We also expect larger salamanders to occupy more aquatic habitat

# FIGURE 8: LINEAR REGRESSION PLOT SHOWING SEAL SALAMANDER BODY SIZE VS COVER OBJECT AREA
```{r}


lr = ggplot(pointsDCHE2, aes(x = covArea, y = mmSVL)) + 
  geom_smooth( method = "lm", col= "#5D695C", alpha = 0.5) + 
  geom_point(aes(col = hab),
             size = 2) + 
  scale_color_manual(values = habCols) + 
  labs(color = "") + 
  xlab(expression(paste("Cover Object Area (", cm^{2}, ")"))) + 
  ylab("Snout-Vent Length (mm)") +
  # labs(title = expression(
  #   paste(
  #     "Cover Object Area vs. ", 
  #     italic("Desmognathus cheaha"), 
  #     " mass"
  #     )
  #   ))+
  
  #facet_wrap(~hab) + 
  theme_bw() + 
  expand_limits(x = 2500) + 
theme(strip.background = element_rect(fill = "transparent", color = "black"), 
      text = element_text(size = 20, family = "Times"), 
       panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA))


lr

```

From this figure, it does look like larger seal salamanders use more aquatic habitat than smaller ones. We will test for a significant relationship here: 

```{r}


habUseDCHE_null = glmmTMB(mmSVL ~ (1|StrmID), data = pointsDCHE2)



habUseDCHE = glmmTMB(mmSVL ~ hab + covArea + (1|StrmID), data = pointsDCHE2)

diagnose(habUseDCHE)


summary(habUseDCHE)

tab_model(habUseDCHE)

habUseDCHE

anova(habUseDCHE_null, habUseDCHE, test = "Chisq")



```



## Are cover area and position autocorrelated?

```{r}
sealStreams = c("STRM01", "STRM02", "STRM04", "STRM06", "STRM07", "STRM08", "STRM10")

pWS = pointsWide %>% filter(StrmID %in% sealStreams & CoverType %in% c("R", "W")) %>%  
  mutate(hab = ifelse(distWater > 50, "terrestrial", NA)) %>%
  mutate(hab = ifelse(cmDepth > 1, "stream", hab)) %>%
  mutate(hab = ifelse(is.na(hab) == T, "margin", hab)) %>%
  mutate(hab = as.factor(hab)) %>% filter(hab %in% c("stream", "margin"))

```

This is available in the text as *Supplemental Table 4*

```{r}


habType = glmmTMB(covArea ~ hab + (1|StrmID), data = pWS)

confint(habType)

summary(habType)

tab_model(habType)

pWS %>% ggplot(aes(x = hab, y = covArea)) + geom_boxplot() + geom_jitter() + 
  theme_bw()


```


It does not look like there's a difference in background points
```{r}

sealPresLabs = c("0" = "Not Occupied by Seal Salamanders", 
                "1" = "Occupied by Seal Salamanders")



pWS %>% filter(hab %in% c("margin", "stream")) %>%
  ggplot(aes(x = hab, y = covArea)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(col = hab), width = 0.1, alpha = 0.5) + 
  scale_color_manual(values = habCols) + 
  ylab(expression(paste("Cover Object Area (", cm^{2}, ")"))) + 
  xlab("Habitat Type") + 
  labs(title = "Comparison of cover object size across habitat type", color = "") + 
  theme_bw() + 
  facet_wrap(~as.factor(sealPres), labeller = as_labeller(sealPresLabs)) + 
  theme(text = element_text(size = 12, family = "Times"), 
        strip.background = element_rect(fill = "transparent", color = "black"))



plots %>% select(Date, PlotCode) %>% arrange(Date)
```












